A tutorial of Bayesian beta regressions with brms in R

### 2025-02-10 manuscript/manuscript.qmd
add example of bad gaussian

Regression models (and extensions like generalised additive models) have become a *de facto* standard for the statistical analysis of a variety of measures in phonetic research [@kirby2018; @politzerahles2018; @tavakoli2024]. However, there is a tendency for researchers to use Gaussian distribution families (i.e. probability distributions for the outcome variable) for any measure that is numeric, irrespective of whether the measure is unbounded, like in truly Gaussian variables, or bounded, like in proportions. A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()` from lme4 [@bates2015], both fit Gaussian regressions by default and the user does not have to specify the distribution family. This tacit default of using Gaussian models is also reflected in teaching practices, where significance test and models using the Gaussian distribution are the first to be taught [@baayen2008; @winter2020], due to their relative simplicity and the fact that regression models with other families are conceptual generalisations of Gaussian regression models.

>>>>

Regression models (and extensions like generalised additive models) have become a *de facto* standard for the statistical analysis of a variety of measures in phonetic research [@kirby2018; @politzerahles2018; @tavakoli2024]. However, there is a tendency for researchers to use Gaussian distribution families (i.e. probability distributions for the outcome variable) for any measure that is numeric, irrespective of whether the measure is unbounded, like in truly Gaussian variables, or bounded, like in proportions. A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()` from lme4 [@bates2015], both fit Gaussian regressions by default and the user does not have to specify the distribution family. This tacit default of using Gaussian models is also reflected in teaching practices, where significance test and models using the Gaussian distribution are the first to be taught [@baayen2008; @winter2020], due to their relative simplicity and the fact that regression models with other families are conceptual generalisations of Gaussian regression models. @fig-bg shows how poorly a Gaussian regression model fits beta-distributed data: the dark blue line is the density of data simulated to be distributed according to a beta distribution; the light blue lines are predicted distributions as generated by a Gaussian model of the data. The predicted Gaussian distributions over-predict values between 0.5 and 0.75 and under-predict values between 0.75 and 1. Moreover, the density peak (median) of the Gaussian distributions is lower than the peak in the original data.

```{r}
#| label: fig-bg
#| fig-cap: "Gaussian model fitted to beta-distributed data. The dark blue line is the density of the beta data, while the light blue lines are simulated densities of the same data based on a Gaussian model."
#| echo: false

bg <- brm(
  x ~ 1,
  data = data.frame(x),
  file = "data/cache/bg"
)

pp_check(bg, ndraws = 10)
```

### 2025-02-10 manuscript/manuscript.tex
add example of bad gaussian

Regression models (and extensions like generalised additive models) have
become a \emph{de facto} standard for the statistical analysis of a
variety of measures in phonetic research
\citep{kirby2018, politzerahles2018, tavakoli2024}. However, there is a
tendency for researchers to use Gaussian distribution families
(i.e.~probability distributions for the outcome variable) for any
measure that is numeric, irrespective of whether the measure is
unbounded, like in truly Gaussian variables, or bounded, like in
proportions. A possible reason is that the base R function for fitting
regression models, \texttt{lm()}, and the lme4 function used to fit
regression models with varying terms, \texttt{lmer()} from lme4
\citep{bates2015}, both fit Gaussian regressions by default and the user
does not have to specify the distribution family. This tacit default of
using Gaussian models is also reflected in teaching practices, where
significance test and models using the Gaussian distribution are the
first to be taught \citep{baayen2008, winter2020}, due to their relative
simplicity and the fact that regression models with other families are
conceptual generalisations of Gaussian regression models.

>>>>

Regression models (and extensions like generalised additive models) have
become a \emph{de facto} standard for the statistical analysis of a
variety of measures in phonetic research
\citep{kirby2018, politzerahles2018, tavakoli2024}. However, there is a
tendency for researchers to use Gaussian distribution families
(i.e.~probability distributions for the outcome variable) for any
measure that is numeric, irrespective of whether the measure is
unbounded, like in truly Gaussian variables, or bounded, like in
proportions. A possible reason is that the base R function for fitting
regression models, \texttt{lm()}, and the lme4 function used to fit
regression models with varying terms, \texttt{lmer()} from lme4
\citep{bates2015}, both fit Gaussian regressions by default and the user
does not have to specify the distribution family. This tacit default of
using Gaussian models is also reflected in teaching practices, where
significance test and models using the Gaussian distribution are the
first to be taught \citep{baayen2008, winter2020}, due to their relative
simplicity and the fact that regression models with other families are
conceptual generalisations of Gaussian regression models.
Figure~\ref{fig-bg} shows how poorly a Gaussian regression model fits
beta-distributed data: the dark blue line is the density of data
simulated to be distributed according to a beta distribution; the light
blue lines are predicted distributions as generated by a Gaussian model
of the data. The predicted Gaussian distributions over-predict values
between 0.5 and 0.75 and under-predict values between 0.75 and 1.
Moreover, the density peak (median) of the Gaussian distributions is
lower than the peak in the original data.

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-bg-1.pdf}}

\caption{\label{fig-bg}Gaussian model fitted to beta-distributed data.
The dark blue line is the density of the beta data, while the light blue
lines are simulated densities of the same data based on a Gaussian
model.}
